\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-02-05 paper.qmd
Add different parameter choices for HARPS data

@@

@@

@@

@@

plot_list2_harps + trimesh_removed_tsne_harps +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))

@@

num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = tSNE1,
                                                cell_area = 1) #43
num_bins_tsne_harps <- 30

@@

@@

num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = tSNE1,
                                                cell_area = 1) #43
#num_bins_tsne_harps <- 30

@@

@@

>>>>

#### Hexagonal grid starting point

Mainly hexagonal grid starts at the lowest x and y values assign for the grid. Another possibility is to shift the grid. That's another interesting parameter to investigate.

hex_full_count_df <- read_rds("data/s_curve/s_curve_hex_8.rds")

shifted_hex_coord_df <- read_rds("data/s_curve/s_curve_hex_8_shifted_hex_coord_df.rds")

ggplot(data = hex_full_count_df, aes(x = c_x, y = c_y)) +
  geom_point(color = "black") +
  geom_point(data = shifted_hex_coord_df, aes(x = c_x, y = c_y), color = "red")

ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
  geom_polygon(fill = NA, color = "#feb24c", aes(group = polygon_id)) +
  geom_polygon(data = hex_full_count_df, aes(x = x, y = y, group = polygon_id),
               fill = NA, color = "#bdbdbd")

## Before shift
ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")

## After shift
ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
```
Compute aic, mse after shifting and compare before and after shifting errors
And investigate how aic, mse change according to the shifted amount

@@

error_df <- read_csv("data/harps/error_harps_tsne.csv", col_names = FALSE)

error_df <- error_df |>
  mutate(X2 = round(X2/100, 2),
         X1 = as.integer(X1)) |>
  arrange(X2)

colnames(error_df) <- c("perplexity", "error (x 100)")

kable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcombharps") |>
  #xtable2kable() |>
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
  row_spec(0, bold = TRUE) |>
  row_spec(3, bold=T, color = "red")
```

@@

num_non_empty_bins_harps <- df_bin_centroids_tsne_harps$hexID |> length()

@@

df_bin_train <- df_bin_tsne_harps
names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])

error_df <- df_all_tsne_harps |>
  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates

  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])

error_df <- error_df |>
  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))

error_df <- error_df |>
  mutate(type = if_else(total <= 0, "no error",
                        if_else(total <= 0.2, "error 0-0.2",
                                if_else(total <= 0.3, "error 0.2-0.3",
                                        if_else(total <= 0.5, "error 0.3-0.5",
                                                if_else(total <= 1, "error 0.5-1",
                                                        "error greter than 1"))))))

plot_list2_harps_error <- error_df |>
    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
    ggplot(aes(x = tSNE1,
               y = tSNE2, color = type))+
    geom_point(alpha=0.5, size = 0.1) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#fee0d2", "#fcbba1",
                                         "#fc9272", "#fb6a4a", "#ef3b2c",
                                         "#cb181d", "#a50f15", "#99000d"))

@@

plot_list2_harps + trimesh_removed_tsne_harps + plot_list2_harps_error +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))

@@

shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
                                                      x = tSNE1, y = tSNE2)

num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)

## To extract bin centroids
hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)

@@

## To find the benchmark value
benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = distance)
benchmark_tsne_harps <- 2.5

@@

shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
                                                      x = tSNE1, y = tSNE2)

num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)

## To extract bin centroids
hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)

@@

## To find the benchmark value
benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = distance)
benchmark_tsne_harps <- 1

@@

df_bin_train <- df_bin_tsne_harps
names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])

error_df <- df_all_tsne_harps |>
  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates

  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])

error_df <- error_df |>
  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))

error_df <- error_df |>
  mutate(type = if_else(total <= 0, "no error",
                        if_else(total <= 0.2, "error 0-0.2",
                                if_else(total <= 0.3, "error 0.2-0.3",
                                        if_else(total <= 0.5, "error 0.3-0.5",
                                                if_else(total <= 1, "error 0.5-1",
                                                        "error greter than 1"))))))

plot_list2_harps_error <- error_df |>
    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
    ggplot(aes(x = tSNE1,
               y = tSNE2, color = type))+
    geom_point(alpha=0.5, size = 0.1) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#fee0d2", "#fcbba1",
                                         "#fc9272", "#fb6a4a", "#ef3b2c",
                                         "#cb181d", "#a50f15", "#99000d"))
```

### 2024-02-05 paper.tex
Add different parameter choices for HARPS data

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi

@@

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, frame hidden, breakable]}{\end{tcolorbox}}\fi

@@

\hypertarget{hexagonal-grid-starting-point}{%
\subsubsection{Hexagonal grid starting
point}\label{hexagonal-grid-starting-point}}

Mainly hexagonal grid starts at the lowest x and y values assign for the
grid. Another possibility is to shift the grid. That's another
interesting parameter to investigate.

\includegraphics{paper_files/figure-pdf/unnamed-chunk-39-1.pdf}

\includegraphics{paper_files/figure-pdf/unnamed-chunk-39-2.pdf}

\includegraphics{paper_files/figure-pdf/unnamed-chunk-39-3.pdf}

\includegraphics{paper_files/figure-pdf/unnamed-chunk-39-4.pdf}

Compute aic, mse after shifting and compare before and after shifting
errors And investigate how aic, mse change according to the shifted
amount

@@

\caption{errors for different parameter combinations}
\centering
\begin{tabular}[t]{r|r}
\hline
\textbf{perplexity} & \textbf{error (x 100)}\\
\hline
\cellcolor{gray!6}{23} & \cellcolor{gray!6}{0.02}\\
\hline
91 & 0.04\\
\hline
\textcolor{red}{\textbf{\cellcolor{gray!6}{40}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{1.05}}}\\
\hline
\end{tabular}
\end{table}
